---
title: "Frequentist+Cross-validation"
author: "Anita Kurm"
date: "December 3, 2018"
output: html_document
---

Let's do some good old frequentist stats. Like here: https://github.com/kurmie/Assignment-2/blob/master/A1_P2_LangASD_trajectories.Rmd

Prepare data:
```{r}
#libraries 
pacman::p_load(ggplot2,dplyr,lme4,MuMIn,lmerTest,gstat,stringr,plyr,caret,modelr,ModelMetrics,Metrics, tidyverse,simr,MASS)

#data
data = read.csv("useformodelling.csv")

#prepare data
data$X = NULL
data$X.1 = NULL
data$rateH=as.numeric(data$rateH)

#square predictors
data$rate2 <-(data$rateH)^2
data$mean2<-(data$meanH)^2
data$EmoPol2<-(data$EmoPol)^2
```

Test hypotheses:

Hypothesis 1: Shortly after a tragic event, public engagement in the gun control debate increases.
Prediction: Positive and small values of temporal proximity to the Stoneman Douglas High shooting will associate with higher rate of relevant to the debate tweets per unit of time

Temporal proximity ~ Tweet Rate per hour + (1|Username)
or
Temporal proximity ~ Tweet Rate per hour^2 + (1|Username)

```{r}
#Model 1
m1 <- lmer(proximity~rateH+(1|Name), data=data)
summary(m1) #output
r.squaredGLMM(m1)
plotm1<-ggplot(data, aes(rateH, proximity))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ x)+
  theme_minimal()+
  xlab("Tweet rate") +
  ylab("Temporal proximity") +
  theme(text = element_text(size=25))+
  ggtitle("Model 1: The line represents the linear effect of tweet rate on proximity")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model1.png", width = 25, height = 25, units = "cm")


m2<- lmer(proximity~1+rateH+I(rateH^2)+(1|Name), data=data) #quadratic
summary(m2)
r.squaredGLMM(m2)
plotm2<-ggplot(data, aes(rateH, proximity))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ poly(x, 2))+
  theme_minimal()+
  xlab("Tweet rate") +
  ylab("Temporal proximity") +
  theme(text = element_text(size=25))+
  ggtitle("Model 2: The line represents the quadratic effect of tweet rate on proximity")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model2.png", width = 25, height = 25, units = "cm")


m3<- lmer(proximity~1+rateH+I(rateH^2)+I(rateH^3)+(1|Name), data=data) #cubic
summary(m3)
r.squaredGLMM(m3)
plotm3<-ggplot(data, aes(rateH, proximity))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ poly(x, 3))+
  theme_minimal()+
  xlab("Tweet rate") +
  ylab("Temporal proximity") +
  theme(text = element_text(size=25))+
  ggtitle("Model 3: The line represents the cubic effect of tweet rate on proximity")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model3.png", width = 25, height = 25, units = "cm")


m4<- lmer(proximity~1+rateH+I(rateH^2)+I(rateH^3)+I(rateH^4)+(1|Name), data=data) #quartic
summary(m4)
r.squaredGLMM(m4)
plotm4<-ggplot(data, aes(rateH, proximity))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ poly(x, 4))+
  theme_minimal()+
  xlab("Tweet rate") +
  ylab("Temporal proximity") +
  theme(text = element_text(size=25))+
  ggtitle("Model 4: The line represents the quartic effect of tweet rate on proximity")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model4.png", width = 25, height = 25, units = "cm")



```

Cubic one looks promising
```{r}
m1inv<- lmer(rateH~1+proximity+(1|Name), data=data) #cubic
summary(m1inv)
r.squaredGLMM(m1inv)
plotm1inv<-ggplot(data, aes(proximity, rateH))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ x)+
  theme_minimal()+
  xlab("Proximity") +
  ylab("Tweet rate") +
  theme(text = element_text(size=25))+
  ggtitle("Model 1 inverted: The line represents the linear effect of proximity on tweet rate")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model1inv.png", width = 25, height = 25, units = "cm")

m2inv<- lmer(rateH~1+proximity+I(proximity^2)+(1|Name), data=data) #cubic
summary(m2inv)
r.squaredGLMM(m2inv)
plotm2inv<-ggplot(data, aes(proximity, rateH))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ poly(x, 2))+
  theme_minimal()+
  xlab("Proximity") +
  ylab("Tweet rate") +
  theme(text = element_text(size=25))+
  ggtitle("Model 2 inverted: The line represents the quadratic effect of proximity on tweet rate")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model2inv.png", width = 25, height = 25, units = "cm")
```


Let's try rescaling variables
```{r}
#new data with scaled variables
ndata=data
ndata$rateH<-scale(ndata$rateH)
ndata$proximity<-scale(ndata$proximity)

#Model 1
m1sc <- lmer(proximity~rateH+(1|Name), data=ndata)
summary(m1sc) #output
r.squaredGLMM(m1sc)
plotm1<-ggplot(ndata, aes(rateH, proximity))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ x)+
  theme_minimal()+
  xlab("Tweet rate") +
  ylab("Temporal proximity") +
  theme(text = element_text(size=25))+
  ggtitle("Model 1 scaled")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model1scaled.png", width = 25, height = 25, units = "cm")


m2sc<- lmer(proximity~1+rateH+I(rateH^2)+(1|Name), data=ndata) #quadratic
summary(m2sc)
r.squaredGLMM(m2sc)
plotm2<-ggplot(ndata, aes(rateH, proximity))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ poly(x, 2))+
  theme_minimal()+
  xlab("Tweet rate") +
  ylab("Temporal proximity") +
  theme(text = element_text(size=25))+
  ggtitle("Model 2 scaled")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model2scaled.png", width = 25, height = 25, units = "cm")

```


Trying out hyperbolic function
```{r}
#(Time~I(1/Requests)) 
#Model 1
m1hyp <- lmer(proximity~I(1/rateH)+(1|Name), data=data)
summary(m1hyp) #output
r.squaredGLMM(m1hyp)
plotm1hyp<-ggplot(data, aes(rateH, proximity))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~I(1/x))+
  theme_minimal()+
  xlab("Tweet rate") +
  ylab("Temporal proximity") +
  theme(text = element_text(size=25))+
  ggtitle("Model 1 hyperbolic")+
  theme(plot.title = element_text(size = 25))+
  theme(legend.key.size = unit(1, "cm"))
ggsave("model1hyp.png", width = 25, height = 25, units = "cm")
```

